---
title: "Survival HW#5"
author: "Tonghao Zhang"
date: "2016/11/14"
output: html_document
---

##12.4

```{r}
library(survival)
library(KMsurv)
library(ggplot2)
data("kidney");attach(kidney)
logfit1<-survreg(Surv(time[type==1],delta[type==1])~1,dist = "loglogistic")
logfit2<-survreg(Surv(time[type==2],delta[type==2])~1,dist = "loglogistic")

mu1<-2.964
sigma1<-0.539
lambda1<-exp(-mu1/sigma1)
alpha1<-1/sigma1
covhatstar1<-logfit1$var
covhat1<-covhatstar1
covhat1[,2]<-covhat1[,2]*logfit1$scale
covhat1[2,]<-covhat1[2,]*logfit1$scale
var.lambda1<-exp(-2*mu1/sigma1)*(covhat1[1,1]/sigma1^2+mu1^2*covhat1[2,2]/sigma1^4-2*mu1*covhat1[1,2]/sigma1^3)
var.alpha1<-covhat1[2,2]/sigma1^4
sd.lambda1<-sqrt(var.lambda1)
sd.alpha1<-sqrt(var.alpha1)

mu2<-5.005
sigma2<-1.74
lambda2<-exp(-mu2/sigma2)
alpha2<-1/sigma2
covhatstar2<-logfit2$var
covhat2<-covhatstar2
covhat2[,2]<-covhat2[,2]*logfit2$scale
covhat2[2,]<-covhat2[2,]*logfit2$scale
var.lambda2<-exp(-2*mu2/sigma2)*(covhat2[1,1]/sigma2^2+mu2^2*covhat2[2,2]/sigma2^4-2*mu2*covhat2[1,2]/sigma2^3)
var.alpha2<-covhat2[2,2]/sigma2^4
sd.lambda2<-sqrt(var.lambda2)
sd.alpha2<-sqrt(var.alpha2)
```
####a)

**group 1**: $\hat(\alpha) = `r alpha1`, \hat(\lambda) = `r lambda1`$, $var(\hat(\alpha)) = `r sd.alpha1`, var(\hat(\lambda)) = `r sd.lambda1`$  

**group 2**: $\hat(\alpha) = `r alpha2`, \hat(\lambda) = `r lambda2`$, $var(\hat(\alpha)) = `r sd.alpha2`, var(\hat(\lambda)) = `r sd.lambda2`$  

####b)

```{r}
expfit1<-survreg(Surv(time,delta)~1,data=kidney[type==1,],dist = "exponential")
expfit2<-survreg(Surv(time,delta)~1,data=kidney[type==2,],dist = "exponential")
LR1<--2*(expfit1$loglik[1]-logfit1$loglik[1])
LR2<--2*(expfit2$loglik[1]-logfit2$loglik[1])
p1L<-1-pchisq(LR1,1)
p2L<-1-pchisq(LR2,1)
wald1<-sigma1*(1/covhat1[2,2])*sigma1
wald2<-sigma2*(1/covhat2[2,2])*sigma1
p1W<-1-pchisq(wald1,1)
p2W<-1-pchisq(wald2,1)
```

**group 1**: likelihood ratio test statistic is `r LR1`, p-value is `r p1L`. Wald test statistic is `r wald1`, p-value $`r p1W`$.

**group 2**: likelihood ratio test statistic is `r LR2`, p-value is `r p2L`. Wald test statistic is `r wald2`, p-value `r p2W`. 

####c)

```{r}
predict(logfit1,data.frame(time=5),type = "quantile",se.fit = TRUE)
predict(logfit2,data.frame(time=5),type = "quantile",se.fit = TRUE)
```

####d)

```{r}
logfit.all<-survreg(Surv(time,delta)~type,data=kidney,dist = "loglogistic")
summary(logfit.all)
```

$H_0$: no effect. p-value = 0.357, fail to reject the null hypothesis.

##12.10

##12.12

##12.14