---
title: "Survival HW#5"
author: "Tonghao Zhang"
date: "2016/11/14"
output: pdf_document
---

##12.4

```{r}
library(survival)
library(KMsurv)
library(ggplot2)
data("kidney");attach(kidney)
logfit1<-survreg(Surv(time[type==1],delta[type==1])~1,dist = "loglogistic")
logfit2<-survreg(Surv(time[type==2],delta[type==2])~1,dist = "loglogistic")

mu1<-2.964
sigma1<-0.539
lambda1<-exp(-mu1/sigma1)
alpha1<-1/sigma1
covhatstar1<-logfit1$var
covhat1<-covhatstar1
covhat1[,2]<-covhat1[,2]*logfit1$scale
covhat1[2,]<-covhat1[2,]*logfit1$scale
var.lambda1<-exp(-2*mu1/sigma1)*(covhat1[1,1]/sigma1^2+mu1^2*covhat1[2,2]/sigma1^4-2*mu1*covhat1[1,2]/sigma1^3)
var.alpha1<-covhat1[2,2]/sigma1^4
sd.lambda1<-sqrt(var.lambda1)
sd.alpha1<-sqrt(var.alpha1)

mu2<-5.005
sigma2<-1.74
lambda2<-exp(-mu2/sigma2)
alpha2<-1/sigma2
covhatstar2<-logfit2$var
covhat2<-covhatstar2
covhat2[,2]<-covhat2[,2]*logfit2$scale
covhat2[2,]<-covhat2[2,]*logfit2$scale
var.lambda2<-exp(-2*mu2/sigma2)*(covhat2[1,1]/sigma2^2+mu2^2*covhat2[2,2]/sigma2^4-2*mu2*covhat2[1,2]/sigma2^3)
var.alpha2<-covhat2[2,2]/sigma2^4
sd.lambda2<-sqrt(var.lambda2)
sd.alpha2<-sqrt(var.alpha2)
```
####a)

**group 1**: $\hat(\alpha) = `r alpha1`, \hat(\lambda) = `r lambda1`$, $var(\hat(\alpha)) = `r sd.alpha1`, var(\hat(\lambda)) = `r sd.lambda1`$  

**group 2**: $\hat(\alpha) = `r alpha2`, \hat(\lambda) = `r lambda2`$, $var(\hat(\alpha)) = `r sd.alpha2`, var(\hat(\lambda)) = `r sd.lambda2`$  

####b)

```{r}
expfit1<-survreg(Surv(time,delta)~1,data=kidney[type==1,],dist = "exponential")
expfit2<-survreg(Surv(time,delta)~1,data=kidney[type==2,],dist = "exponential")
LR1<--2*(expfit1$loglik[1]-logfit1$loglik[1])
LR2<--2*(expfit2$loglik[1]-logfit2$loglik[1])
p1L<-1-pchisq(LR1,1)
p2L<-1-pchisq(LR2,1)
wald1<-sigma1*(1/covhat1[2,2])*sigma1
wald2<-sigma2*(1/covhat2[2,2])*sigma1
p1W<-1-pchisq(wald1,1)
p2W<-1-pchisq(wald2,1)
```

**group 1**: likelihood ratio test statistic is `r LR1`, p-value is `r p1L`. Wald test statistic is `r wald1`, p-value $`r p1W`$.

**group 2**: likelihood ratio test statistic is `r LR2`, p-value is `r p2L`. Wald test statistic is `r wald2`, p-value `r p2W`. 

####c)

```{r}
S1<-1/(1+lambda1*5^alpha1)
S2<-1/(1+lambda2*5^alpha2)
kpfit1<-survfit(Surv(time,delta)~1,data = kidney[type==1,])
kpfit1.est<-kpfit1$surv[3];CI.1<-c(kpfit1$lower[3],kpfit1$upper[3])
kpfit2<-survfit(Surv(time,delta)~1,data = kidney[type==2,])
kpfit2.est<-kpfit2$surv[3];CI.2<-c(kpfit2$lower[3],kpfit2$upper[3])
```

The point estimate of first group is `r S1` [0.5285844,1], while KP estimate is `r kpfit1.est`, [`r CI.1`].

The point estimate of group type 2 is `r S2` [0.727708,1], while KP estimate is `r kpfit2.est`, [`r CI.2`].

####d)

```{r}
logfit.all<-survreg(Surv(time,delta)~as.factor(type),data=kidney,dist = "loglogistic")
mu.all<-2.9590
sigma.all<-1.06
gamma.all<-0.4588
alpha.all<-1/sigma.all
lambda.all<-exp(-mu.all/sigma.all)
beta.all<--gamma.all/sigma.all

afactor<-exp(beta.all)^(1/alpha.all)
podd<-exp(-beta.all)
```

$H_0$: no effect. p-value = 0.357, fail to reject the null hypothesis.

The point estimate of relative odds is `r podd`, point estimate of acceleration factor is `r afactor`.

##12.10

```{r,fig.height=6}
library(ggplot2)
library(gridExtra)
cumhaz.1<--log(kpfit1$surv);cumhaz.2<--log(kpfit2$surv)
expfit1<-survreg(Surv(time,delta)~1,data = kidney[type==1,],dist = "exp")
expfit2<-survreg(Surv(time,delta)~1,data = kidney[type==2,],dist = "exp")
lambda1.exp<-exp(-3.48)
lambda2.exp<-exp(-4.01)
p1<-qplot(kpfit1$time,cumhaz.1)+labs(x="",y="")+ggtitle("exponetial")+geom_line()
p2<-qplot(kpfit2$time,cumhaz.2)+labs(x="",y="")+ggtitle("exponetial")+geom_line()

weibfit1<-survreg(Surv(time,delta)~1,data = kidney[type==1,],dist = "weibull")
mu1.weib<-3.194;sigma1.weib<-0.627
lambda1.weib<-exp(-mu1.weib/sigma1.weib);alpha1.weib<-1/sigma1.weib
p3<-qplot(log(kpfit1$time),log(cumhaz.1))+labs(x="",y="")+ggtitle("weibull")+geom_line()
weibfit2<-survreg(Surv(time,delta)~1,data = kidney[type==2,],dist = "weibull")
mu2.weib<-5.411;sigma2.weib<-1.85
lambda2.weib<-exp(-mu2.weib/sigma2.weib);alpha2.weib2<-1/sigma2.weib
p4<-qplot(log(kpfit2$time),log(cumhaz.2))+labs(x="",y="")+ggtitle("weibull")+geom_line()

lognormal1<-survreg(Surv(time,delta)~1,data = kidney[type==1,],dist = "loggaussian")
lognormal2<-survreg(Surv(time,delta)~1,data = kidney[type==2,],dist = "loggaussian")
p5<-qplot(log(kpfit1$time),qnorm(1-exp(-cumhaz.1)))+labs(x="",y="")+ggtitle("log-normal")+geom_line()
p6<-qplot(log(kpfit2$time),qnorm(1-exp(-cumhaz.2)))+labs(x="",y="")+ggtitle("log-normal")+geom_line()
p7<-qplot(log(kpfit1$time),log(exp(-log(kpfit1$surv))-1))+labs(x="",y="")+ggtitle("log-logistic")+geom_line()
p8<-qplot(log(kpfit2$time),log(exp(-log(kpfit2$surv))-1))+labs(x="",y="")+ggtitle("log-logistic")+geom_line()
grid.arrange(p1,p2,p3,p4,ncol=2)
```
```{r,fig.height=6}
grid.arrange(p5,p6,p7,p8,ncol=2)
```

As there not much observed event in type 2 group, its estimate is more like stepwise, not smooth. From the plots above, it seems to me weibull and log-logistic are better models.

##12.12

```{r,fig.height=3,fig.width=3}
tfit2<-c(0.5,2.5,3.5,6.5,15.5)
tfit1<-c(3.5,4.5,5.5,5.5,8.5)
qplot(tfit1,tfit2)
```

Again number of observed events in group 2 is too small, I can only find 5 quantiles. From this five points qqplot log-logistic model is ok. 

##12.14

####a)
```{r,fig.height=3,fig.width=3}
data("hodg")
G<-1*(hodg$gtype==1);D<-1*(hodg$dtype==1)
weibfit<-survreg(Surv(time,delta)~G+D+G*D,data = hodg,dist = "weibull")
summary(weibfit)
mu<-6.952;sigma<-1.65;gamma<-c(-3.319,-1.160,5.358)
beta<--gamma/sigma
lambda<-exp(-mu/sigma);alpha<-1/sigma

coxsnell.weib<-lambda*exp(beta[1]*G+beta[2]*D+beta[3]*G*D)*hodg$time^alpha
kp.cs<-survfit(Surv(coxsnell.weib,hodg$delta)~1)
qplot(kp.cs$time,-log(kp.cs$surv))
```

There is a severe departure from straight line, which indicates Weibull is not a suitable model here.

####b)

```{r,fig.height=3,fig.width=3}
mres.weib<-hodg$delta-coxsnell.weib
dres.weib<-sign(mres.weib)*(-2*(mres.weib+hodg$delta*log(hodg$delta-mres.weib)))^0.5
qplot(hodg$time,dres.weib)
```

Deviance plot indicates Weibull model is not suitable, it is very far from what we expect from a random noise plot.

####c)

```{r,fig.height=3,fig.width=3}
loglogistic<-survreg(Surv(time,delta)~G+D+G*D,data = hodg,dist = "loglogistic")
mu<--5.991;sigma<-1.26;gamma<-c(-2.874,-0.786,4.411)
beta<--gamma/sigma
lambda<-exp(-mu/sigma);alpha<-1/sigma

#coxsnell.ll<-log(1+exp(-loglogistic$linear.predictors*alpha)*hodg$time^alpha)
coxsnell.ll<-log(1/(1+lambda*exp(beta[1]*G+beta[2]*D+beta[3]*G*D)*hodg$time^alpha))
kp.cs.ll<-survfit(Surv(coxsnell.ll,hodg$delta)~1)
qplot(kp.cs.ll$time,-log(kp.cs.ll$surv))
mres.ll<-hodg$delta-coxsnell.ll
dres.ll<-resid(loglogistic,type = "deviance")
qplot(hodg$time,dres.ll)
```

Log-logistic model is not suitable neither.